dispRity manual

### 2017-08-04 doc/dispRity-manual.Rmd
Fixed typos etc to 'Metric run through'

Of course, functions can be more complex and involve multiple operations such as the `centroids` function (see `?dispRity.metric`) that calculates the Euclidean distance between each elements and the a position in the ordinated space.
The `make.metric` function implemented in `dispRity` is designed to help testing and finding the dimension-level of the functions.
This function tests:

1.  if your function can deal with a `matrix` or a `vector` as an input;
2.  which is your function's dimension-level according to it's output (dimension-level 1, 2 or 3, see above);
3.  whether the function can properly be implemented in the `dispRity` function (the function is fed into a `lapply` loop).

@@

```{r}
## Which dimension-level is the mean function? And is able to be used in the dispRity?
make.metric(mean)

## Same questions from the prod.rows function:
make.metric(prod.rows)

## Same questions for the var function:
make.metric(var)
```

A silent version of the function is also available to be included in your code.
This can be done using the option `silent = TRUE` and will simply output the dimension-level of the metric.

@@

### metric argument in `dispRity`

@@

```{r}
## Measuring disparity as the standard deviation of all the value of the
## ordinated matrix (dimension-level 1 function).
summary(dispRity(BeckLee_mat50, metric = sd))

@@

## Same but using a different functions order for the metric argument
disparity2 <- summary(dispRity(BeckLee_mat50, metric = c(variances, sd, var)),
                      round = 10)

@@

In these examples, we considered disparity to be a single value.
For example, in the previous one, we defined disparity as the standard deviation of the variances of each column of the variance/covariance matrix (`metric = c(variances, sd, var)`).
It is however possible to calculate [disparity as a distribution](#disparity-as-a-distribution).

@@

Several disparity metrics are directly implemented in the `dispRity` package.
The detailed list can be found in `?dispRity.metric` along with some explanations on each metric.

@@

1: This function uses an estimation of the eigen value that only works for MDS or PCoA ordinations (not PCA).

2: Note that by default, the centroid is the centroid of the elements.
It can however be fixed to an arbitrary value by using the `centroid` argument `centroids(space, centroid = rep(0, ncol(space)))`, for example the origin of the ordinated space.

### Equations and implementations:
Some of the functions described below are directly implemented in the `dispRity` package and do not require any dependencies  ([see implementation here](https://github.com/TGuillerme/dispRity/blob/master/R/dispRity.metric.R)).

\begin{equation}
    ellipse.volume = \frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} (\lambda_{i}^{0.5})
\end{equation}
With *k* being the number of dimensions, $\lambda_i$ is the eigen value of each dimension.

\begin{equation}
    diagonal = \sqrt{\sum_{i=1}^{k}|max(k_i) - min(k_i)|}
\end{equation}
With *k* being the dimensions.

\begin{equation}
    ranges = |max(k_i) - min(k_i)|
\end{equation}
With *k* being the dimensions.

\begin{equation}
    variances = \sigma^{2}{k_i}
\end{equation}
With *k* being the dimensions and $\sigma^{2}$ their variance.

\begin{equation}
    centroids = \sqrt{\sum_{i=1}^{n}{({k}_{n}-Centroid_{k})^2}}
\end{equation}
With *n* being each element in the ordinated space, *k* being its dimensions and $Centroid_{k}$ their mean (or can be set to an arbitrary value) and.

>>>>

Of course, functions can be more complex and involve multiple operations such as the `centroids` function (see `?dispRity.metric`) that calculates the Euclidean distance between each element and the centroid of the ordinated space.
The `make.metric` function implemented in `dispRity` is designed to help test and find the dimension-level of the functions.
This function tests:

1.  If your function can deal with a `matrix` or a `vector` as an input;
2.  Your function's dimension-level according to its output (dimension-level 1, 2 or 3, see above);
3.  Whether the function can be implemented in the `dispRity` function (the function is fed into a `lapply` loop).

@@

```{r}
## Which dimension-level is the mean function? And can it be used in dispRity?
make.metric(mean)

## Which dimension-level is the prod.rows function? And can it be used in dispRity?
make.metric(prod.rows)

## Which dimension-level is the var function? And can it be used in dispRity?
make.metric(var)
```

A non verbose version of the function is also available.
This can be done using the option `silent = TRUE` and will simply output the dimension-level of the metric.

<!-- NC: My terminology changes may have broken this bit -->

@@

### Metrics in the `dispRity` function

@@

```{r}
## Measuring disparity as the standard deviation of all the values of the
## ordinated matrix (dimension-level 1 function).
summary(dispRity(BeckLee_mat50, metric = sd))

@@

## Same as above but using a different function order for the metric argument
disparity2 <- summary(dispRity(BeckLee_mat50, metric = c(variances, sd, var)),
                      round = 10)

@@

In these examples, we considered disparity to be a single value.
For example, in the previous example, we defined disparity as the standard deviation of the variances of each column of the variance/covariance matrix (`metric = c(variances, sd, var)`).
It is, however, possible to calculate [disparity as a distribution](#disparity-as-a-distribution).

@@

Several disparity metrics are implemented in the `dispRity` package.
The detailed list can be found in `?dispRity.metric` along with some description of each metric.

@@

1: This function uses an estimation of the eigenvalue that only works for MDS or PCoA ordinations (not PCA).

2: Note that by default, the centroid is the centroid of the elements.
It can, however, be fixed to a different value by using the `centroid` argument `centroids(space, centroid = rep(0, ncol(space)))`, for example the origin of the ordinated space.

### Equations and implementations
Some of the functions described below are implemented in the `dispRity` package and do not require any other packages to calculate ([see implementation here](https://github.com/TGuillerme/dispRity/blob/master/R/dispRity.metric.R)).

\begin{equation}
    ellipse.volume = \frac{\pi^{k/2}}{\Gamma(\frac{k}{2}+1)}\displaystyle\prod_{i=1}^{k} (\lambda_{i}^{0.5})
\end{equation}
Where *k* is the number of dimensions, and $\lambda_i$ is the eigenvalue of each dimension.

\begin{equation}
    diagonal = \sqrt{\sum_{i=1}^{k}|max(k_i) - min(k_i)|}
\end{equation}
Where *k* is the number of dimensions.

\begin{equation}
    ranges = |max(k_i) - min(k_i)|
\end{equation}
Where *k* is the number of dimensions.

\begin{equation}
    variances = \sigma^{2}{k_i}
\end{equation}
Where *k* is the number of dimensions, and $\sigma^{2}$ is their variance.

\begin{equation}
    centroids = \sqrt{\sum_{i=1}^{n}{({k}_{n}-Centroid_{k})^2}}
\end{equation}
Where *n* is each element in the ordinated space, *k* is the number of dimensions, and $Centroid_{k}$ is their mean (or can be set to another value).
